Thesis outline

### 2017-06-06 KellyDouglass/Thesis/Thesis_sections/GV.tex
BPT distribution of abundances, Appendix section on physics of transitions added

During my estimation of the chemical content of galaxies with $M_r > -20$, I 
note that the oxygen abundances of galaxies classified as having an AGN by their 
emission line ratios in the BPT diagrams exhibit higher nitrogen abundances and 
normal oxygen abundance values when compared to star-forming galaxies of a 
similar stellar mass and large-scale environment.  The same abundance signature 
is seen in the the oxygen and nitrogen abundance results seen in Fig. 
\ref{fig:Z_hist} of the green valley galaxies.  This might indicate a 
link between green valley galaxies and a host AGN.  This would support other 
studies of the transition of galaxies from the blue sequence to the red cloud, 
which attribute the quenching of star formation to accretion of a central black 
hole \citep[e.g.,][]{Croton06,Stasinska08}.

>>>>

\begin{figure}
    \includegraphics[width=0.49\textwidth]{Images/GV/1sig_all_BPT_t3_12logOHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{Images/GV/1sig_all_BPT_t3_12logNHrelations_dust_hist}
    \includegraphics[width=0.49\textwidth]{Images/GV/1sig_all_BPT_t3_logNOrelations_dust_hist}
    \caption[Gas-phase chemical abundance distributions for star-forming, 
    composite, and AGN galaxies]{Distribution of the gas-phase oxygen abundance 
    (O/H, top left), nitrogen abundance (N/H, top right), and N/O ratio (bottom 
    center) of the SDSS DR7 galaxies with detectable emission lines for 
    abundance calculations with the Direct $T_e$ method.  Star-forming galaxies 
    are shown with a solid cyan line, while composite galaxies are a dashed 
    orange line and galaxies with an AGN are represented as a dotted magenta 
    line.  Composite galaxies have lower oxygen abundances than both 
    star-forming galaxies and those with an AGN.  The nitrogen abundance range 
    is different for each of the three classes, with star-forming galaxies 
    having the lowest nitrogen abundances, AGN having the highest, and 
    composite galaxies in the middle.  Therefore, star-forming galaxies have 
    N/O ratios $< 0$, while composite galaxies and galaxies with an AGN have 
    N/O ratios between -1 and 0.5.}
    \label{fig:BPT_abund}
\end{figure}

To understand the gas-phase chemical abundances of the green valley galaxies, we 
compare Fig. \ref{fig:Z_hist} with the distribution of the gas-phase chemical 
abundances of star-forming and composite galaxies and galaxies with an AGN shown 
in Fig. \ref{fig:BPT_abund}.  We see that the galaxies with an AGN have the most 
similar distributions in the three abundance ratios as the green valley 
galaxies: both groups have similar oxygen abundances (O/H) and higher nitrogen 
abundances (N/H) and N/O ratios than the star-forming galaxies.  This might 
indicate a link between green valley galaxies and a host AGN.  The presence of 
an AGN in green valley galaxies would support other studies of the transition of 
galaxies from the blue sequence to the red cloud, which attribute the quenching 
of star formation to accretion of a central black hole \citep[e.g.,][]{Croton06,
Stasinska08}.

### 2017-06-06 KellyDouglass/Thesis/Thesis_sections/Paper1.tex
BPT distribution of abundances, Appendix section on physics of transitions added

%-------------------------------------------------------------------------------
% Oxygen
\subsection{[O{\sc iii}]}\label{sec:O3}

@@

%-------------------------------------------------------------------------------
\subsection{[O{\sc ii}]}
A less temperature-sensitive line than [O{\sc iii}] $\lambda 4363$, the 
[O{\sc ii}] $\lambda 3727$ forbidden transition doublet of singly-ionized 
oxygen is often used in metallicity calculations.  With an electron temperature 
and density, its flux provides an estimate of the abundance of singly-ionized 
oxygen.  In SDSS spectra, this line can be observed for objects with a redshift 
greater than 0.02.  However, because dwarf galaxies are inherently faint objects 
($M_r > -17$), they are targeted for spectroscopy in SDSS only out to redshift 
$z\sim 0.03$, thus we can only estimate the metallicity of dwarf galaxies in the 
redshift range $0.02 < z < 0.03$.

%-------------------------------------------------------------------------------
% Sulfur
\subsection{[S{\sc ii}]}
Just as we are able to measure the electron temperature from [O{\sc iii}] 
transitions, we can estimate the electron number density from [S{\sc ii}] 
transitions.  Below a density of about $100 \text{ cm}^{-3}$, the [S{\sc ii}] 
$\lambda 6716 / \lambda 6731$ ratio has a weak dependence on the density.  All 
our galaxies fall within this low-density regime, so we assume a low-density 
limit of $n_e = 100 \text{ cm}^{-3}$.

@@

%-------------------------------------------------------------------------------
\subsection{Direct $T_e$ method}
We use the method published by \cite{Izotov06}, which is based on the 
astrophysics in \cite{Osterbrock89}.  It makes use of the [O{\sc iii}] 
$\lambda 4363$, $\lambda \lambda 4959, 5007$ lines and the [O{\sc ii}] 
$\lambda 3727$ doublet.  While often regarded as the most accurate estimate of 
the metallicity, it is difficult to employ due to the restrictions on 
[O{\sc iii}] $\lambda 4363$.  Consequently, this method is best suited for 
low-redshift, low-metallicity galaxies.  The electron temperature is derived by 
solving the following system of equations:
\begin{equation}
	t_3 = \frac{1.432}{\log[(\lambda 4959 + \lambda 5007)/\lambda 4363] - \log C_T}
\end{equation}
where $t_3 = 10^{-4} T_e(\text{O}^{++})$ and
\begin{equation}
	C_T = (8.44 - 1.09t_3 + 0.5t_3^2 - 0.08t_3^3)\frac{1 + 0.0004x_3}{1 + 0.044x_3}
\end{equation}
where $x_3 = 10^{-4} n_e t_3^{-0.5}$.  The ionic abundances are then found with 
the equations
\begin{align}
	12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) &= \log \frac{\lambda 3727}{\text{H}\beta} + 5.961 + \frac{1.676}{t_2} - 0.40\log t_2 - 0.034t_2 + \log (1+1.35x_2) \\
	12 + \log \left( \frac{\text{O}^{++}}{\text{H}^+} \right) &= \log \frac{\lambda 4959 + \lambda 5007}{\text{H}\beta} + 6.200 + \frac{1.251}{t_3} - 0.55\log t_3 - 0.014t_3
\end{align}
where $t_2 = 10^{-4} T_e(\text{O}^+)$, $t_2 = 0.7t_3 + 0.3$ \citep{Garnett92}, 
and $x_2 = 10^{-4} n_e t_2^{-0.5}$.  \cite{Andrews13} show that this relation 
between $T_e(\text{O}^+)$ and $T_e(\text{O}^{++})$ may overestimate the 
temperature in the low ionization zone, causing the calculated metallicities to 
be underestimated.  Because we care only about the relative metallicity values 
of the galaxies, this effect will only affect our results in galaxies where 
O$^+$ dominates the oxygen abundance (where O$^+$/O$^{++} > 1$) in higher 
temperature regions (or low metallicities).  As shown in Fig. 
\ref{fig:O2O3_ratio}, this affects perhaps fifteen galaxies and does not change 
our conclusions.

@@

\begin{table}
    \centering
    \begin{tabular}{cccc}
        \hline \\
         & $Z < 7.6$ & $7.6 \leq Z < 8.2$ & $Z \geq 8.2$\\
        \hline \\
        \hline \\
        \multicolumn{4}{c}{$1 \sigma$ restriction on [O{\sc iii}] $\lambda 4363$}\\
        \hline \\
        Void & 9.52\% (4)   & 66.67\% (28) & 23.81\% (10)\\
        Wall & 19.10\% (17) & 59.55\% (53) & 21.35\% (19)\\
        \hline \\
        \multicolumn{4}{c}{No restriction on [O{\sc iii}] $\lambda 4363$}\\
        \hline \\
        Void & 13.33\% (16) & 45.83\% (55)  & 40.83\% (49)\\
        Wall & 23.26\% (60) & 46.12\% (119) & 30.62\% (79)\\
        \hline
    \end{tabular}
    \caption[Metallicity distribution of 135 dwarf galaxy sample]{Percentages of 
    galaxies with calculated metallicities within the labeled metallicity 
    ranges, with the number of galaxies in each category in parentheses.  
    Removing the S/N restriction on [O{\sc iii}] $\lambda 4363$ especially 
    increases the number of dwarf galaxies with high metallicities, changing the 
    distribution so that void dwarf galaxies have higher metallicities than wall 
    dwarf galaxies.  However, due to the large uncertainties in the metallicity 
    estimates for $12 + \log (\text{O}/\text{H}) > 8.2$, this difference in the 
    distributions may not be statistically significant.\label{tab:Percents}}
\end{table}

>>>>

%-------------------------------------------------------------------------------
% Oxygen
\subsection{[\ion{O}{3}]}\label{sec:O3}

@@

%-------------------------------------------------------------------------------
\subsection{[\ion{O}{2}]}

A less temperature-sensitive line than [O{\sc iii}] $\lambda 4363$, the 
[O{\sc ii}] $\lambda 3727$ forbidden transition doublet of singly-ionized 
oxygen is often used in metallicity calculations.  With an electron temperature 
and density, its flux provides an estimate of the abundance of singly-ionized 
oxygen.  In SDSS spectra, this line can be observed for objects with a redshift 
greater than 0.02.  However, because dwarf galaxies are inherently faint objects 
($M_r > -17$), they are targeted for spectroscopy in SDSS only out to redshift 
$z\sim 0.03$, thus we can only estimate the metallicity of dwarf galaxies in the 
redshift range $0.02 < z < 0.03$.

%-------------------------------------------------------------------------------
% Sulfur
\subsection{[\ion{S}{2}]}

Just as we are able to measure the electron temperature from [O{\sc iii}] 
transitions, we can estimate the electron number density from [S{\sc ii}] 
transitions.  Below a density of about $100 \text{ cm}^{-3}$, the [S{\sc ii}] 
$\lambda 6716 / \lambda 6731$ ratio has a weak dependence on the density.  All 
our galaxies fall within this low-density regime, so we assume a low-density 
limit of $n_e = 100 \text{ cm}^{-3}$.

@@

%-------------------------------------------------------------------------------
\subsection{Direct $T_e$ method}

We use the method published by \cite{Izotov06}, which is based on the 
astrophysics in \cite{Osterbrock89}.  It makes use of the [O{\sc iii}] 
$\lambda 4363$, $\lambda \lambda 4959, 5007$ lines and the [O{\sc ii}] 
$\lambda 3727$ doublet.  While often regarded as the most accurate estimate of 
the metallicity, it is difficult to employ due to the restrictions on 
[O{\sc iii}] $\lambda 4363$.  Consequently, this method is best suited for 
low-redshift, low-metallicity galaxies.  The electron temperature is derived by 
solving the following system of equations:
\begin{equation}
	t_3 = \frac{1.432}{\log[(\lambda 4959 + \lambda 5007)/\lambda 4363] - \log C_T}
\end{equation}
where $t_3 = 10^{-4} T_e(\text{O}^{++})$ and
\begin{equation}
	C_T = (8.44 - 1.09t_3 + 0.5t_3^2 - 0.08t_3^3)\frac{1 + 0.0004x_3}{1 + 0.044x_3}
\end{equation}
where $x_3 = 10^{-4} n_e t_3^{-0.5}$.  The ionic abundances are then found with 
the equations
\begin{align}
	12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) &= \log \frac{\lambda 3727}{\text{H}\beta} + 5.961 + \frac{1.676}{t_2} - 0.40\log t_2 - 0.034t_2 + \log (1+1.35x_2) \\
	12 + \log \left( \frac{\text{O}^{++}}{\text{H}^+} \right) &= \log \frac{\lambda 4959 + \lambda 5007}{\text{H}\beta} + 6.200 + \frac{1.251}{t_3} - 0.55\log t_3 - 0.014t_3
\end{align}
where $t_2 = 10^{-4} T_e(\text{O}^+)$, $t_2 = 0.7t_3 + 0.3$ \citep{Garnett92}, 
and $x_2 = 10^{-4} n_e t_2^{-0.5}$.  \cite{Andrews13} show that this relation 
between $T_e(\text{O}^+)$ and $T_e(\text{O}^{++})$ may overestimate the 
temperature in the low ionization zone, causing the calculated metallicities to 
be underestimated.  Because we care only about the relative metallicity values 
of the galaxies, this effect will only affect our results in galaxies where 
O$^+$ dominates the oxygen abundance (where O$^+$/O$^{++} > 1$) in higher 
temperature regions (or low metallicities).  As shown in Fig. 
\ref{fig:O2O3_ratio}, this affects perhaps fifteen galaxies and does not change 
our conclusions.

@@

\begin{table}
    \centering
    \begin{tabular}{cccc}
        \hline
         & $Z < 7.6$ & $7.6 \leq Z < 8.2$ & $Z \geq 8.2$\\
        \hline
        \hline
        \multicolumn{4}{c}{$1 \sigma$ restriction on [O{\sc iii}] $\lambda 4363$}\\
        \hline
        Void & 9.52\% (4)   & 66.67\% (28) & 23.81\% (10)\\
        Wall & 19.10\% (17) & 59.55\% (53) & 21.35\% (19)\\
        \hline
        \multicolumn{4}{c}{No restriction on [O{\sc iii}] $\lambda 4363$}\\
        \hline
        Void & 13.33\% (16) & 45.83\% (55)  & 40.83\% (49)\\
        Wall & 23.26\% (60) & 46.12\% (119) & 30.62\% (79)\\
        \hline
    \end{tabular}
    \caption[Metallicity distribution of 135 dwarf galaxy sample]{Percentages of 
    galaxies with calculated metallicities within the labeled metallicity 
    ranges, with the number of galaxies in each category in parentheses.  
    Removing the S/N restriction on [O{\sc iii}] $\lambda 4363$ especially 
    increases the number of dwarf galaxies with high metallicities, changing the 
    distribution so that void dwarf galaxies have higher metallicities than wall 
    dwarf galaxies.  However, due to the large uncertainties in the metallicity 
    estimates for $12 + \log (\text{O}/\text{H}) > 8.2$, this difference in the 
    distributions may not be statistically significant.\label{tab:Percents}}
\end{table}

### 2017-06-06 KellyDouglass/Thesis/Thesis_sections/Paper2.tex
BPT distribution of abundances, Appendix section on physics of transitions added

% ------------------------------------------------------------------------------
\subsection{[N{\sc ii}]}

The energy-level diagram for the various transitions of [N{\sc ii}] is very 
similar to that of [O{\sc iii}], since they have the same electron ground-state 
configuration ((1s)$^2$(2s)$^2$(2p)$^2$).  The similarities can be seen in Fig. 
\ref{fig:transitions_P2}.  Therefore, an estimate of the electron temperature can 
be made from the [N{\sc ii}] $\lambda 5755$ emission line.  However, this line 
is weaker than the [O{\sc iii}] $\lambda 4363$ auroral line (since there is less 
N than O in galaxies), so we use the [O{\sc iii}] auroral line for our 
temperature estimates, as in Paper I.  After obtaining a temperature and density 
estimate, we use the [N{\sc ii}] $\lambda \lambda 6548, 6584$ doublet to 
estimate the abundance of singly ionized nitrogen in a galaxy.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper2/NIIOIII_energy_level_diagram-figure0}
    \caption[{[\ion{O}{3}] and [\ion{N}{2}] energy-level diagram}]{Energy-level 
    diagram for [O{\sc iii}] and [N{\sc ii}] $(2p^2)$ ions.  The most important 
    transitions are shown; all are in the visible spectrum.  These forbidden 
    transitions in both oxygen and nitrogen provide an estimate of the electron 
    temperature in the interstellar gas.  Because oxygen is more abundant, we 
    use the oxygen lines to estimate the temperature of the gas.}
    \label{fig:transitions_P2}
\end{figure}

@@

We use the same method to calculate the nitrogen abundance as in Paper I to 
estimate the oxygen abundance.  However, here we use the [N{\sc ii}] 
$\lambda \lambda 6548, 6584$ doublet instead of the [O{\sc ii}] $\lambda 3727$ 
and [O{\sc iii}] $\lambda \lambda 4959,5007$ doublets.  Because the temperature 
estimate depends on the auroral line [O{\sc iii}] $\lambda 4363$, this method is 
often difficult to employ.  As a result, it works best with low-redshift, 
low-metallicity galaxies.  The electron temperature is derived by solving the 
following system of equations:
\begin{equation}
	t_3 = \frac{1.432}{\log[(\lambda 4959 + \lambda 5007)/\lambda 4363] - \log C_T}
\end{equation}
where $t_3 = 10^{-4} T_e(\text{O}^{++})$ and
\begin{equation}
	C_T = (8.44 - 1.09t_3 + 0.5t_3^2 - 0.08t_3^3)\frac{1 + 0.0004x_3}{1 + 0.044x_3}
\end{equation}
where $x_3 = 10^{-4} n_e t_3^{-0.5}$.  The ionic abundances are then found with 
the equations
\begin{align}
    12 + \log \left( \frac{\text{O}^{++}}{\text{H}^+} \right) &= \log \frac{\lambda 4959 + \lambda 5007}{\text{H}\beta} + 6.200 + \frac{1.251}{t_3} - 0.55\log t_3 - 0.014t_3 \\
	12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) &= \log \frac{\lambda 3727}{\text{H}\beta} + 5.961 + \frac{1.676}{t_2} - 0.40\log t_2 - 0.034t_2 + \log (1+1.35x_2) \\
	12 + \log \left( \frac{\text{N}^+}{\text{H}^+} \right) &= \log \frac{\lambda 6548 + \lambda 6584}{\text{H}\beta} + 6.234 + \frac{0.950}{t_2} - 0.42\log t_2 - 0.027t_2 + \log (1 + 0.116x_2)
\end{align}
where $t_2 = 10^{-4} T_e(\text{O}^+)$ and $x_2 = 10^{-4} n_e t_2^{-0.5}$.  We 
assume that $T_e(\text{N}^+) = T_e(\text{O}^+)$.

>>>>

% ------------------------------------------------------------------------------
\subsection{[\ion{N}{2}]}

The energy-level diagram for the various transitions of [\ion{N}{2}] is very 
similar to that of [\ion{O}{3}], since they have the same electron ground-state 
configuration ((1s)$^2$(2s)$^2$(2p)$^2$).  The similarities can be seen in Fig. 
\ref{fig:transitions_P2}.  Therefore, an estimate of the electron temperature can 
be made from the [\ion{N}{2}] $\lambda 5755$ emission line.  However, this line 
is weaker than the [\ion{O}{3}] $\lambda 4363$ auroral line (since there is less 
N than O in galaxies), so we use the [\ion{O}{3}] auroral line for our 
temperature estimates, as in Paper I.  After obtaining a temperature and density 
estimate, we use the [\ion{N}{2}] $\lambda \lambda 6548, 6584$ doublet to 
estimate the abundance of singly ionized nitrogen in a galaxy.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Images/Paper2/NIIOIII_energy_level_diagram-figure0}
    \caption[{[\ion{O}{3}] and [\ion{N}{2}] energy-level diagram}]{Energy-level 
    diagram for [\ion{O}{3}] and [\ion{N}{2}] $(2p^2)$ ions.  The most important 
    transitions are shown; all are in the visible spectrum.  These forbidden 
    transitions in both oxygen and nitrogen provide an estimate of the electron 
    temperature in the interstellar gas.  Because oxygen is more abundant, we 
    use the oxygen lines to estimate the temperature of the gas.}
    \label{fig:transitions_P2}
\end{figure}

@@

We use the same method to calculate the nitrogen abundance as in Paper I to 
estimate the oxygen abundance.  However, here we use the [\ion{N}{2}] 
$\lambda \lambda 6548, 6584$ doublet instead of the [\ion{O}{2}] $\lambda 3727$ 
and [\ion{O}{3}] $\lambda \lambda 4959,5007$ doublets.  Because the temperature 
estimate depends on the auroral line [\ion{O}{3}] $\lambda 4363$, this method is 
often difficult to employ.  As a result, it works best with low-redshift, 
low-metallicity galaxies.  The electron temperature is derived by solving the 
following system of equations:
\begin{equation}
	t_3 = \frac{1.432}{\log[(\lambda 4959 + \lambda 5007)/\lambda 4363] - \log C_T}
\end{equation}
where $t_3 = 10^{-4} T_e(\text{O}^{++})$ and
\begin{equation}
	C_T = (8.44 - 1.09t_3 + 0.5t_3^2 - 0.08t_3^3)\frac{1 + 0.0004x_3}{1 + 0.044x_3}
\end{equation}
where $x_3 = 10^{-4} n_e t_3^{-0.5}$.  The ionic abundances are then found with 
the equations
\begin{align}
    12 + \log \left( \frac{\text{O}^{++}}{\text{H}^+} \right) &= \log \frac{\lambda 4959 + \lambda 5007}{\text{H}\beta} + 6.200 + \frac{1.251}{t_3} - 0.55\log t_3 - 0.014t_3 \\
	12 + \log \left( \frac{\text{O}^+}{\text{H}^+} \right) &= \log \frac{\lambda 3727}{\text{H}\beta} + 5.961 + \frac{1.676}{t_2} - 0.40\log t_2 - 0.034t_2 + \log (1+1.35x_2) \\
	12 + \log \left( \frac{\text{N}^+}{\text{H}^+} \right) &= \log \frac{\lambda 6548 + \lambda 6584}{\text{H}\beta} + 6.234 + \frac{0.950}{t_2} - 0.42\log t_2 - 0.027t_2 + \log (1 + 0.116x_2)
\end{align}
where $t_2 = 10^{-4} T_e(\text{O}^+)$ and $x_2 = 10^{-4} n_e t_2^{-0.5}$.  We 
assume that $T_e(\text{N}^+) = T_e(\text{O}^+)$.
